# Preparation

```{r}
library("devtools")
library("tidyterra")
library("ggplot2")
library("prism")
library("stringr")
library("terra")
library("data.table")
library("dplyr")
library("sf")
library("stars")

devtools::load_all()
```

# Generate datasets 

## fairway grid

```{r}
fairway_grid <-
  sf::st_read("data-raw/fwy05_grid_NAD.shp") %>%
  st_transform(4326)

usethis::use_data(fairway_grid, overwrite = TRUE)
```

## World

```{r}
library("rnaturalearth")
library("rnaturalearthdata")

world <-
  ne_countries(scale = "medium", returnclass = "sf") %>%
  dplyr::select(name)

usethis::use_data(world, overwrite = TRUE)
```

## Nebraska wells as sf

```{r}
data(wells_ne)

wells_ne_sf <-
  sf::st_as_sf(
    wells_ne,
    coords = c(
      "longdd",
      "latdd"
    ),
    crs = 4269
  ) %>%
  dplyr::mutate(gw_extracted = runif(nrow(.), min = 50, max = 200))

usethis::use_data(wells_ne_sf, overwrite = TRUE)
```

## High-Plains aquifer boundary

```{r}
hp_boundary <-
  sf::st_read("data-raw/hp_bound2010.shp") %>%
  dplyr::filter(AQUIFER == "High Plains aquifer") %>%
  dplyr::select(geometry) %>%
  st_transform(4269)

usethis::use_data(hp_boundary, overwrite = TRUE)
```

## Nebraska county boundaries

```{r}
ne_counties <-
  tigris::counties(state = "NE", cb = TRUE, progress_bar = FALSE) %>%
  dplyr::select(STATEFP, COUNTYFP, NAME) %>%
  data.table::setnames(names(.), tolower(names(.)))

sf::st_crs(ne_counties)$wkt <- gsub("°|º", "", sf::st_crs(ne_counties)$wkt)

usethis::use_data(ne_counties, overwrite = TRUE)
```

## Colorado county boundaries

```{r}
co_counties <-
  tigris::counties(state = "CO", cb = TRUE, progress_bar = FALSE) %>%
  dplyr::select(STATEFP, COUNTYFP, NAME) %>%
  data.table::setnames(names(.), tolower(names(.)))

sf::st_crs(co_counties)$wkt <- gsub("°|º", "", sf::st_crs(co_counties)$wkt)

usethis::use_data(co_counties, overwrite = TRUE)
```

## Railroads in Nebraska

(https://catalog.data.gov/dataset/tiger-line-shapefile-2019-nation-u-s-rails-national-shapefile)

```{r}
download.file("http://www2.census.gov/geo/tiger/TIGER2019/RAILS/tl_2019_us_rails.zip", destfile = "data-raw/us_rails.zip")

railroads_ne <-
  sf::st_read("data-raw/us_rails/tl_2019_us_rails.shp") %>%
  .[ne_counties, ] %>%
  dplyr::select(LINEARID)

usethis::use_data(railroads_ne, overwrite = TRUE)
```

## NE corn harvested acres

```{r}
get_quickstat_data <- function(year, data_item) {
  desc <-
    tidyUSDA::allDataItem %>%
    grep(pattern = data_item, ., value = TRUE) %>%
    .[!grepl(., pattern = "NET")]

  temp_data <- tidyUSDA::getQuickstat(
    key = keyring::key_get("usda-nass"),
    program = "SURVEY",
    data_item = desc,
    geographic_level = "COUNTY",
    year = year,
    geometry = TRUE
  ) %>%
    data.table() %>%
    .[, .(year, county_code, state_fips_code, short_desc, Value, geometry)]

  return(temp_data)
}

corn_acres_ne <-
  lapply(
    as.character(2020:2023),
    \(x) get_quickstat_data(x, "CORN, GRAIN - ACRES HARVESTED")
  ) %>%
  rbindlist() %>%
  setnames("Value", "acre") %>%
  .[, short_desc := NULL] %>%
  .[state_fips_code == 31, ] %>%
  .[county_code != "998", ] %>%
  .[, state_fips_code := NULL] %>%
  st_as_sf() %>%
  st_simplify(dTolerance = 100)

usethis::use_data(corn_acres_ne, overwrite = TRUE)
```

## IA corn nitrogen

This is a made-up corn nitrogen use by county.

```{r}
set.seed(893434)

ia_nitrogen <-
  tigris::counties(state = "IA") %>%
  setnames(names(.), tolower(names(.))) %>%
  dplyr::select(countyfp) %>%
  dplyr::mutate(nitrogen_rate := runif(nrow(.), min = 100, max = 300))

usethis::use_data(ia_nitrogen, overwrite = TRUE)
```

## Hydrologic units

```{r}
ia_bbox <-
  st_bbox(ia_nitrogen) %>%
  st_as_sfc()

huc_ia <-
  st_read("data-raw/huc250k.shp") %>%
  st_transform(st_crs(ia_bbox)) %>%
  .[ia_bbox, op = st_within] %>%
  st_simplify(dTolerance = 100) %>%
  dplyr::select(HUC_CODE)

usethis::use_data(huc_ia, overwrite = TRUE)
```

## Yield and as-applied nitrogen

```{r}
soy_yield <-
  st_read("data-raw/yield-simple1.shp") %>%
  .[seq(1, nrow(.), by = 10), ] %>%
  dplyr::select("Yld_Vol_Dr") %>%
  rename(yield = Yld_Vol_Dr) %>%
  dplyr::mutate(yield_id = 1:nrow(.)) %>%
  st_transform_utm()

#--- remove non-ASCII character ---#
sf::st_crs(soy_yield)$wkt <- gsub("°|º", "", st_crs(soy_yield)$wkt)

usethis::use_data(soy_yield, overwrite = TRUE)

as_applied_s_rate <-
  st_read("data-raw/as-applied-simple1.shp") %>%
  .[seq(1, nrow(.), by = 10), ] %>%
  dplyr::select("Rt_Apd_Ct_") %>%
  rename(seed_rate = Rt_Apd_Ct_) %>%
  dplyr::mutate(seed_id = 1:nrow(.)) %>%
  st_transform_utm()

#--- remove non-ASCII character ---#
sf::st_crs(as_applied_s_rate)$wkt <- gsub("°|º", "", st_crs(as_applied_s_rate)$wkt)

usethis::use_data(as_applied_s_rate, overwrite = TRUE)

# ggplot() +
#   geom_sf(data = yield, size = 0.4, color = "red") +
#   geom_sf(data = as_applied_n_rate, size = 0.4, color = "blue")
```

# Basic sfg

## Points
```{r}
#--- create points ---#
point_1 <- st_point(c(2, 2))
point_2 <- st_point(c(1, 1))
point_3 <- st_point(c(1, 3))

#--- combine the points to make a single  sf of points ---#
points <- list(point_1, point_2, point_3) %>%
  st_sfc() %>%
  st_as_sf() %>%
  mutate(point_name = c("point 1", "point 2", "point 3"))

usethis::use_data(points, overwrite = TRUE)
```

## Lines

```{r}
#--- create lines ---#
line_1 <- st_linestring(rbind(c(0, 0), c(2.5, 0.5)))
line_2 <- st_linestring(rbind(c(1.5, 0.5), c(2.5, 2)))

#--- combine the points to make a single  sf of points ---#
lines <- list(line_1, line_2) %>%
  st_sfc() %>%
  st_as_sf() %>%
  mutate(line_name = c("line 1", "line 2"))

usethis::use_data(lines, overwrite = TRUE)
```

## Polygons

```{r}
#--- create polygons ---#
polygon_1 <- st_polygon(list(
  rbind(c(0, 0), c(2, 0), c(2, 2), c(0, 2), c(0, 0))
))

polygon_2 <- st_polygon(list(
  rbind(c(0.5, 1.5), c(0.5, 3.5), c(2.5, 3.5), c(2.5, 1.5), c(0.5, 1.5))
))

polygon_3 <- st_polygon(list(
  rbind(c(0.5, 2.5), c(0.5, 3.2), c(2.3, 3.2), c(2, 2), c(0.5, 2.5))
))

#--- combine the polygons to make an sf of polygons ---#
polygons <- list(polygon_1, polygon_2, polygon_3) %>%
  st_sfc() %>%
  st_as_sf() %>%
  mutate(polygon_name = c("polygon 1", "polygon 2", "polygon 3"))

usethis::use_data(polygons, overwrite = TRUE)
```

# Raster data

`SpatRaster` object cannot be saved as *.rda as it saves a pointer. So, they are converted to `stars` object and saved.

## Satellite image, blue, green, and red

```{r}
#--- blue ---#
reflec_blue <-
  terra::rast("data-raw/Annos_062718_RE_transparent_reflectance_blue.tif") %>%
  aggregate(fact = 10)

set.names(reflec_blue, "blue")

# terra::writeRaster(reflec_blue, "data-files-for-participants/reflec_blue.tif")

reflec_blue <- stars::st_as_stars(reflec_blue)
sf::st_crs(reflec_blue)$wkt <- gsub("°|º", "", st_crs(reflec_blue)$wkt)

usethis::use_data(reflec_blue, overwrite = TRUE)

#--- red ---#
reflec_red <-
  terra::rast("data-raw/Annos_062718_RE_transparent_reflectance_red.tif") %>%
  aggregate(fact = 10)

set.names(reflec_red, "red")

# terra::writeRaster(reflec_red, "data-files-for-participants/reflec_red.tif")

reflec_red <- stars::st_as_stars(reflec_red)
sf::st_crs(reflec_red)$wkt <- gsub("°|º", "", st_crs(reflec_red)$wkt)

usethis::use_data(reflec_red, overwrite = TRUE)

#--- green ---#
reflec_green <-
  terra::rast("data-raw/Annos_062718_RE_transparent_reflectance_green.tif") %>%
  aggregate(fact = 10)

set.names(reflec_green, "green")

# terra::writeRaster(reflec_green, "data-files-for-participants/reflec_green.tif")

reflec_green <- stars::st_as_stars(reflec_green)
sf::st_crs(reflec_green)$wkt <- gsub("°|º", "", st_crs(reflec_green)$wkt)

usethis::use_data(reflec_green, overwrite = TRUE)
```

## NIR and RED

```{r}
#--- NIR ---#
NIR <-
  terra::rast("data-raw/NIR.tif") %>%
  aggregate(fact = 20)

set.names(NIR, "NIR")

NIR <- stars::st_as_stars(NIR)
sf::st_crs(NIR)$wkt <- gsub("°|º", "", st_crs(NIR)$wkt)

usethis::use_data(NIR, overwrite = TRUE)

#--- RED ---#
RED <-
  terra::rast("data-raw/RED.tif") %>%
  aggregate(fact = 20)

set.names(RED, "RED")

RED <- stars::st_as_stars(RED)
sf::st_crs(RED)$wkt <- gsub("°|º", "", st_crs(RED)$wkt)

usethis::use_data(RED, overwrite = TRUE)
```

## PRISM for NE

```{r}
library(lubridate)
options(prism.path = "data-raw")

prism::get_prism_dailys(
  type = "ppt",
  minDate = "2012-08-01",
  maxDate = "2012-08-05",
  keepZip = FALSE
)

all_prism_files <-
  list.dirs("data-raw") %>%
  .[str_detect(., "PRISM")] %>%
  .[str_detect(., "ppt")] %>%
  lapply(
    .,
    \(x) {
      list.files(x, full.names = TRUE) %>%
        .[str_detect(., "\\.bil")] %>%
        .[!str_detect(., "aux")]
    }
  ) %>%
  unlist()

prism_2012_aug <-
  rast(all_prism_files) %>%
  terra::crop(ne_counties) %>%
  terra::mask(ne_counties)

names(prism_2012_aug) <- seq(ymd("2012-08-01"), ymd("2012-08-05"), by = "day")

prism_2012_aug <- stars::st_as_stars(prism_2012_aug)

usethis::use_data(prism_2012_aug, overwrite = TRUE)
```

### PRISM practice data

Single day PRISM data:

```{r}
options(prism.path = "data-files-for-participants")

prism::get_prism_dailys(
  type = "ppt",
  minDate = "2012-08-01",
  maxDate = "2012-08-01",
  keepZip = FALSE
)
```

### PRISM exercise data 

Single day PRISM data simplified:

```{r}
prism_us <-
  terra::rast("data-raw/PRISM_ppt_stable_4kmD2_20120801_bil/PRISM_ppt_stable_4kmD2_20120801_bil.bil") %>%
  aggregate(fact = 5)

prism_us <- stars::st_as_stars(prism_us)

usethis::use_data(prism_us, overwrite = TRUE)
```

### PRISM splitted to practice terra::merge

```{r}
options(prism.path = "data-raw")

prism::get_prism_dailys(
  type = "tmax",
  minDate = "2012-08-01",
  maxDate = "2012-08-01",
  keepZip = FALSE
)

prism_us <-
  terra::rast("data-raw/PRISM_tmax_stable_4kmD2_20120801_bil/PRISM_tmax_stable_4kmD2_20120801_bil.bil")

ne_counties_sd <-
  tigris::counties(state = "NE", cb = TRUE) %>%
  dplyr::filter(NAME %in% c("Douglas", "Saunders"))

prism_saunders <- terra::crop(prism_us, dplyr::filter(ne_counties_sd, NAME == "Saunders"))
prism_douglas <- terra::crop(prism_us, dplyr::filter(ne_counties_sd, NAME == "Douglas"))

# ggplot() +
#   geom_spatraster(data = prism_saunders) +
#   geom_spatraster(data = prism_douglas)

# merged <- terra::merge(prism_saunders, prism_douglas)

prism_saunders <- stars::st_as_stars(prism_saunders)
prism_douglas <- stars::st_as_stars(prism_douglas)

usethis::use_data(prism_saunders, overwrite = TRUE)
usethis::use_data(prism_douglas, overwrite = TRUE)
```

```{r}
values(prism_saunders) <- 90
merge(prism_douglas, prism_saunders) %>% plot()
merge(prism_saunders, prism_douglas) %>% plot()
```


## gridMET

```{r}
downloader::download(
  url = str_c("http://www.northwestknowledge.net/metdata/data/pr_2018.nc"),
  destfile = "data-raw/pr_2018.nc",
  mode = "wb"
)

pr_2018_gm <- terra::rast("data-raw/pr_2018.nc")

#--- first ten days ---#
gm_precip_2018 <- subset(pr_2018_gm, 1:10)

#--- write as an nc file ---#
terra::writeCDF(gm_precip_2018, "data-files-for-participants/gm_precip_2018.nc", overwrite = TRUE)

# terra::writeRaster(gm_precip_2018, "data-files-for-participants/gm_precip_2018.nc", overwrite = TRUE)
```

# On-farm nitrogen experiment

## Treatment blocks

```{r}
treatment_blocks <-
  st_read("data-raw/ZP_TRT_Blocks_Buff.shp") %>%
  dplyr::select(Product, Tgt_Rate_l, Replicatio) %>%
  setnames(names(.), c("product", "target_rate", "replication", "geometry"))

usethis::use_data(treatment_blocks, overwrite = TRUE)
```

## Fake corn yield

```{r}
corn_yield <-
  st_read("data-raw/Stevens_Centuro_YLD_shp.shp") %>%
  #--- transform the CRS so that it is consistent with the raster data ---#
  st_transform(st_crs(treatment_blocks)) %>%
  #--- use every 100 observations ---#
  .[st_as_sfc(st_bbox(treatment_blocks)), ] %>%
  #--- create fake yield data ---#
  mutate(yield = runif(n(), min = 50, max = 240))

usethis::use_data(corn_yield, overwrite = TRUE)

# ggplot() +
#   geom_sf(data = treatment_blocks) +
#   geom_sf(data = yield)
```

## NDRE

```{r}
NDRE <-
  terra::rast("data-raw/NDRE.tif") %>%
  aggregate(fact = 20) %>%
  stars::st_as_stars()

usethis::use_data(NDRE, overwrite = TRUE)

# ggplot() +
#   geom_sf(data = treatment_blocks) +
#   geom_sf(data = yield) +
#   geom_spatraster(data = rast(NDRE))

```

## County yield data

```{r}
county_yield <-
  readRDS("data-raw/county_yield.rds") %>%
  .[, .[, .(corn_yield, soy_yield, year, county_code, state_name, d0_5_9, d1_5_9, d2_5_9, d3_5_9, d4_5_9, geometry)]] %>%
  .[!is.na(corn_yield), ] %>%
  .[state_name %in% c("Nebraska", "Colorado", "Kansas"), ] %>%
  as_tibble()

usethis::use_data(county_yield, overwrite = TRUE)
```

